Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation.}

### 2020-07-15 journal paper/journal_main.tex
Moved some macro to the main file.
Minor edits in "theory/extensions".
Changed an appendix title.
DEG -> BEG, DG -> BG
Minor edits.
Removed "in progress" from a section title.
Revised some macro.
fix wording in abstract

@@

% notation
%%% advanced notation
\newcommand{\OPT}{\term{OPT}}
\newcommand{\rew}{\term{rew}}  % Bayesian-expected reward after n local rounds
\newcommand{\REP}{\term{REP}} % posterior mean reward
\newcommand{\PMR}{\term{PMR}} % posterior mean reward
\newcommand{\support}{\term{support}}

\newcommand{\BIR}{\term{BIR}} % Bayesian Instantaneous Regret
\newcommand{\regret}{R^{\term{inst}}} % regret
\newcommand{\regretWC}{\regret_{\term{wc}}} % regret

@@

% algorithms
\newcommand{\StaticGreedy}{\term{StaticGreedy}}
\newcommand{\DynGreedy}{\term{BayesianGreedy}}
\newcommand{\DynamicGreedy}{\DynGreedy}
\newcommand{\Thompson}{\term{ThompsonSampling}}

@@

We consider a stylized duopoly model in which two firms face the same instance of multi-armed bandits. Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen. We study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for consumers. We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low consumer welfare. However, we find that weakening competition by providing firms with some ``free" consumers incentivizes better exploration strategies and increases consumer welfare.  We investigate two channels for weakening the competition: relaxing the rationality of consumers and giving one firm a first-mover advantage. We provide a mix of theoretical results and numerical simulations.

@@

\newpage
\section{Theoretical results: the \TheoryModel}
\label{sec:theory}
\subfile{content/sec-theory}

@@

\section{Background for non-specialists: the three classes of algorithms [IN PROGRESS]}
\label{sec:related-classes}
\subfile{content/sec-bg}

>>>>

% in connection with "a little greedy goes a long way"
%     stuff for the "greedy modification".
\newcommand{\gr}{\texttt{gr}}
\newcommand{\BIRgr}{\BIR^{\gr}}
\newcommand{\alggr}{\alg[\gr]}
\newcommand{\rewgr}{\rew_{\gr}}

@@

% notation
%%% advanced notation
\newcommand{\OPT}{\term{OPT}}
\newcommand{\rew}{\term{rew}}  % Bayesian-expected reward after n local rounds
\newcommand{\PMR}{\term{PMR}} % posterior mean reward
\newcommand{\support}{\term{support}}

\newcommand{\BIR}{\term{BIR}} % Bayesian Instantaneous Regret

@@

% algorithms
\newcommand{\StaticGreedy}{\term{StaticGreedy}}
\newcommand{\DynGreedy}{\term{BayesianGreedy}}
\newcommand{\DynamicGreedy}{\DynGreedy}
\newcommand{\DynamicEpsGreedy}{\term{BayesianEpsilonGreedy}}
\newcommand{\Thompson}{\term{ThompsonSampling}}

@@

We consider a stylized duopoly model in which two firms face the same \gaedit{multi-armed bandit instance.} Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen. We study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for consumers. We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low consumer welfare. However, we find that weakening competition by providing firms with some ``free" consumers incentivizes better exploration strategies and increases consumer welfare.  We investigate two channels for weakening the competition: relaxing the rationality of consumers and giving one firm a first-mover advantage. We provide a mix of theoretical results and numerical simulations.

@@

\newpage
\section{Theoretical results: the \TheoryModel}
\label{sec:theory}
\subfile{content/sec-theory}
\newpage

@@

\section{Background for non-specialists: multi-armed bandits}
\label{sec:related-classes}
\subfile{content/sec-bg}
